Assignment 1
========================================================
## Reading the data file
```{r}
 dataFile <- read.csv('activity.csv')
```
## Daily steps taken
### Histogram of the total number of steps taken each day
```{r}
dailySteps <- aggregate(dataFile$steps,list(dataFile$date),sum,na.rm=TRUE)

barplot(dailySteps$x, names.arg=as.Date(dailySteps$Group.1), ylab="Steps", xlab="Days")
```
### Daily mean steps
```{r}
dailyMeanSteps <- aggregate(dataFile$steps,list(dataFile$date),mean,na.rm=TRUE)
colnames(dailyMeanSteps) <- c('date','steps')
dailyMeanSteps <- replace(dailyMeanSteps,is.na(dailyMeanSteps),0)
dailyMeanSteps
```

### Daily median steps

```{r}
medianFrame <- data.frame(date = character(0),median_steps=numeric(0))
aaa <- numeric()
dateVector <- as.vector(dataFile$date)
stepsVector <- as.vector(dataFile$steps)

for(i in seq_along(dateVector)){    
    if(i==length(dateVector) || dateVector[i] != dateVector[i+1]){        
        medianFrame <- rbind(medianFrame, data.frame(date = dateVector[i],median_steps = median(aaa)))
        aaa = numeric()
    }  
    else if(!is.na(stepsVector[i])){
            aaa <- c(aaa,stepsVector[i])           
    }
}
medianFrame
```

### Average steps per interval
```{r}
intervalAvgSum <- 0
intervalVector <- as.vector(dataFile$interval)
avgDataFrame <- data.frame(interval=numeric(0),avg_steps=numeric(0))
intervalAvg <- 0
intervalMaxAvg <- -1
maxAvgInterval <- 1
for(i in seq_along(stepsVector)){
         if(is.na(stepsVector[i])){
              intervalAvgSum <- intervalAvgSum+0
        }else{
              intervalAvgSum <- intervalAvgSum+stepsVector[i]
          }
       intervalAvg <-intervalAvgSum/i  
       avgDataFrame <- rbind(avgDataFrame, data.frame(interval=intervalVector[i],avg_steps=intervalAvg))     
       
       if(intervalAvg > intervalMaxAvg){
          intervalMaxAvg <- intervalAvg
          maxAvgInterval <- intervalVector[i]
       }
}
plot(avgDataFrame$interval,avgDataFrame$avg_steps,type="l")

```
### The interval containing the maximum steps on average is `r maxAvgInterval` with value equal to `r intervalMaxAvg`

## Missing values

```{r}
missingValueRowCount <- 0
for(i in seq_along(stepsVector)){
    if(is.na(intervalVector[i]) || is.na(stepsVector[i]) || is.na(dateVector[i])){
        missingValueRowCount <- missingValueRowCount+1
    }
}
```
### There are `r missingValueRowCount` rows with missing(NA) values  

### As such missing values are only present in steps column we desided to replace them with average of mean steps per interval, see code below
```{r}
intervalMeanSteps <- aggregate(dataFile$steps,list(dataFile$interval),mean,na.rm=TRUE)
colnames(intervalMeanSteps) <- c('interval','steps')
intervalMeanSteps <- replace(intervalMeanSteps,is.na(intervalMeanSteps),0)
intervalMeanStepsVector <- as.vector(intervalMeanSteps$steps)
intervalMeanAverage <- 0
intervalMeanSum <- 0
for(i in seq_along(intervalMeanStepsVector)){
          if(!is.na(intervalMeanStepsVector[i])){
                intervalMeanSum <- intervalMeanSum+intervalMeanStepsVector[i]
              }
      }
  intervalMeanAverage <- intervalMeannSum/length(intervalMeanStepsVector)
  intervalMeanAverage

for(i in seq_along(stepsVector)){
  if(is.na(stepsVector[i])){
      stepsVector[i] <- intervalMeanAverage
  }
}
```
### Creating a new dataset with the following code: 

```{r}
newDataSet <- data.frame(steps<-stepsVector,date<-dateVector,interval<-intervalVector)  
```

### Making a histogram of the total number of steps taken each day and calculate and report mean and median total number of steps taken per day according to the new data set

1. Histogram of the total number of steps taken each day
```{r}
newDailySteps <- aggregate(newDataSet$steps,list(newDataSet$date),sum,na.rm=TRUE)

barplot(newDailySteps$x, names.arg=as.Date(newDailySteps$Group.1), ylab="Steps", xlab="Days")
```

2. Daily mean steps
```{r}
newDailyMeanSteps <- aggregate(newDataSet$steps,list(newDataSet$date),mean,na.rm=TRUE)
colnames(newDailyMeanSteps) <- c('date','steps')
newDailyMeanSteps <- replace(newDailyMeanSteps,is.na(newDailyMeanSteps),0)
newDailyMeanSteps
```

3. Daily median steps

```{r}
newMedianFrame <- data.frame(date = character(0),median_steps=numeric(0))
bbb <- numeric()
newDateVector <- as.vector(newDataSet$date)
newStepsVector <- as.vector(newDataSet$steps)

for(i in seq_along(newDateVector)){    
    if(i==length(newDateVector) || newDateVector[i] != newDateVector[i+1]){        
        newMedianFrame <- rbind(newMedianFrame, data.frame(date = newDateVector[i],median_steps = median(bbb)))
        bbb = numeric()
    }  
    else if(!is.na(newStepsVector[i])){
            bbb <- c(bbb,newStepsVector[i])           
    }
}
newMedianFrame
```
### These values differ a bit from the estimates from the first part of the assignment.